---
title: "CS 422: Homework #1"
author: "Tiffany Wong, Illinois Institute of Technology"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
--- 

### Part 1-A

```{r} 
# import dplyr package 
# install.packages("dplyr") 
library("dplyr") 
```

```{r} 
# read in the college.csv into a dataframe 
df <- read.csv("/Users/tiffwong/Desktop/cs422/hw/hw 1/College.csv") 

# take rows 0 to 5 and then columns 1,5,8,10 from the df 
print.data.frame(df[0:5, c(1, 5, 8, 10)])
```

### Part 1-B

```{r}
# counts number of private colleges WITHOUT looping 
private <- length(which(df$Private == "Yes"))

# counts number of public colleges WITHOUT looping 
public <- length(which(df$Private == "No"))

#print answer statement
cat("There are", private, "private colleges, and", public, "public colleges in the dataset.") 
```

### Part 1-C 

```{r}
# create subset of df with the following columns: Private, Apps, Accept, Enroll, PhD, perc.alumni, S.F.Ratio, and Grad.Rate 
new_df <- df %>% select(Name, Private, Apps, Accept, Enroll, PhD, perc.alumni, S.F.Ratio, Grad.Rate) 
new_df 

```

### Part 1-D(i) +  (iii) extra credit
```{r}
# subset the df by private colleges first
private_df <- new_df[(new_df$Private == 'Yes'), ]

# create a vector for the number of bins in the histogram for private phd holders
colors = rainbow(10)

# histogram of PhD holders in private colleges 
hist(private_df$PhD, main='Number of PhD Holders in Private Colleges', 
     xlab='Number of PhD Holders', col=colors) 

``` 

### Part 1-D(ii) +  (iii) extra credit

```{r}
# subset the df by private colleges first
public_df <- new_df[(new_df$Private == 'No'), ]

# create a vector for the number of bins in the histogram for public phd holders
colors = rainbow(8)

# histogram of PhD holders in public colleges 
hist(public_df$PhD, main='Number of PhD Holders in Public Colleges', 
     xlab='Number of PhD Holders', col=colors) 

``` 

### Part 1-E(i) 
```{r} 
# order the dataframe in ascending order to get the lowest to highest grad rate 
ascending_order <- new_df[order(new_df$Grad.Rate), ]

print.data.frame(ascending_order[1:5, c('Name', 'Grad.Rate')]) 
```

### Part 1-E(ii) 

```{r} 
# order the dataframe in descending order to get the highest to lowest grad rate 
descending_order <- new_df[order(-new_df$Grad.Rate), ]

print.data.frame(descending_order[1:5, c('Name', 'Grad.Rate')]) 
```

### Part 1-F 

```{r} 
# install.packages("psych") 
library(psych) 
```

### Part 1-I(i)

```{r} 
# using on pairs.panels() to create correlation plot with three columns: phd, s.f.ratio, grad rate 
pairs.panels(new_df[ , c('PhD', 'S.F.Ratio', 'Grad.Rate')])  

```

### Part 1-I(ii) 

Which two attributes have the highest correlation? Explain the reason behind this positive correlation (i.e., does this correlation make sense? Why?). 

The two attributes with the highest correlation is Grad.Rate and PhD, meaning there is a high correlation between the graduation rate of a college with the number of PhD holders in said school. This makes sense to me because it's possible that a lot of colleges have a pipeline where students who graduated from their undergrad program go straight into their Masters/PhD programs. So that would make sense why the higher the graduation rate is, the more PhD holders in the school. 


### Part 1-I(iii) 

Which two attributes exhibit the lowest correlation? Explain the reason behind this negative correlation (i.e., does this correlation make sense? Why?).

The two attributes with the lowest correlation is Grad.Rate and S.F.Ratio, meaning there is a high correlation between the graduation rate of a college with the student to faculty ratio in said school. This correlation could be higher in order for it to make sense to me because I think having a lower student to professor ratio (meaning there aren't that many students in a class with one professor), could be beneficial to helping a student guarantee graduation. However, the variable is labeled student to faculty, meaning it could be including staff members whose job isn't to help the students academically. All in all, I think it makes sense that the two variables have the lowest correlation because I personally won't attribute a college's graduation rate to its student to faculty ratio. 


### Part 1-G 

Which alumni donate more to their colleges --- those who go to public schools or those who go to private schools? 

As I found out from the boxplot visual below, alumni donate more to their college if they went to to a private school. 

```{r} 
# using private_df, public_df, and boxplot(), answer the following question with boxplots to compare: 
# Which alumni donate more to their colleges --- those who go to public schools or those who go to private schools? 

boxplot(perc.alumni~Private, data=df, main='Alumni Donations vs Private and Public Colleges', xlab='Private Status', ylab='Percentage Alumni Donates', col='purple', border='pink') 

``` 

### Part 1-H(i) 

What is the median expenditure per student?
Looking at the graph created in 1-H(iii) below, the median expenditure per student looks to be about $7,500, where the 50% is above 5k expenditure, but not quite 10k. 


### Part 1-H(ii) 

80% of the students pay less than how many dollars?
80% of students pay less than $15,000. 



### Part 1-H(iii)
```{r} 
plot(ecdf(df$Expend), main="Student's Expenditure Cumulative Distribution Function Plot", xlab='Expenditure Amount', ylab='Cumulative Distribution Function Amount') 
grid(nx=20, ny=20, col='blue')  
``` 








